---
title: "voorbeeld oscillatie analyse"
author: "Steven Wink"
date: "7 december 2016"
output: html_document
---

 function that fits fourier transform or spline basis functions using the fda package toolkit and extracts parameters of interest for in vitro cell biologists
 data input: 
  `Frame`: integer index of time series (frame numbers)  
  `value`: measurement associated with tracked cell that oscillates.
  `locationID`: string uniquely identifying location and track number of time lapse location and track number, which should hold unique track identifiers you perform the function over. 
 `plotFit`: plot your fits, only do this for a subset during calibration, because writing the pdfs takes alot of time
 `nbasis`: number of basis functions, increase till smooth fit is obtained. At a certain point no improvement will be noticed.
 `lambda`: smoothing factor for data. 0 is no smoothing (default max(Frame)/50000) ) 
 `THpeaks`: threshold value ratio of pred_peak - lowess(f) and median value, smaller values will not be defined as peaks. (basicly the fitted peaks should be above the lowess regression line)   
 `f`: fraction of data to be used for lowess smoothing, lowess is used for THpeaks. recommend 0.4-0.5. For noisy data increase (noise not id as peak). If lowess line follows fit too closely no maximal/ minima will be defined (also dependent on THpeaks.)
 `timeInterval`: time between consecutive frames, format "%H:%M:%S"
 `aboveRegr` : number of points that should be above regression line to identify as peak area) - depcited in plots as thick black line
 `basisType` : fourier or spline, fourier for osicillatory data and spline for non oscillatory
 `slopeDomain` : vector of two, first entry the start time - second entry the end time. format "%H:%M:%S. Average slope in this domain is returned
 `timeToTH`: vector of two, first entry the value, second entry the type: 'relative' or 'absoluut'. returns either the time that the relative threshold value is reached or the time the absoluut value is reached. relative threshold: a fraction between 0 and 1 (0, 1) is defined as the maximum fitted value of the corresponding track * timeToTH. absoluut threshold is defined as the time the timeToTH value is reached
`signal1` and `signal2`: extra data of same length as value and Frame, if provided mean is calculated (usefull for quality control). 
`signal1` can also be used together with `signal1TH` to return the time the `signal1TH` value is reached.
`signal1TH`: threshold value : returns the time this value in `signal1` is reached.
`slopeTH`: regions defined as low slope: first derive below value is defined as low slope. Low slope values are used to determine peak areas (together with local maxima). slopeTH increase will make yellow line gaps larger
`xCoord` and `yCoord`: vectors same length as Frame and value, if provided the tracks will be plotted.
`pix.x`and `pix.y`: resolution of images, needed for plotting tracks. Origin is top left.



the plots: 
a raw fitting plot with residual standard error  
a plot that visualizes the influence of the various tuning paramters. black circles: raw data of track, Blue circle: peak, red circle: valley, yellow dots (zoom in): non-low-slope regions, black dots:regions where fit > lowess regression, black solid-fit line: the model fit, vertical black broken lines: on each side of such a line a defined peak area - but only if it contains a blue circle. These areas are used to determine per - peak measurements. 


## TODO: nuclear and cytosol time courses stipjes in oscillatie  
## TODO: validation with tracknumbers  
## solution: x en y - coordinaten meegeven als argument, als meegegeven - dan mee plotten met domain plots.

```{r data_prep }
require(reshape2)
require(ggplot2)
require(fda)
require(data.table)



datainput <- fread( file = "reorderedTrackData20160107.txt", sep ="\t")


datainput$variable <- gsub("TP_", "", datainput$variable)
datainput$variable <- as.numeric(as.character(datainput$variable))


datainput <- as.data.table(datainput)


test <- data.frame(colA = 1:100, colB = 1:100, variable = sample(c("een", "twee"), 100, replace=T), value = runif(100))
testw <- dcast( colA + colB  ~ variable , data = test)

unique(datainput[ , .id])

datainput <- datainput[.id %in% c('obj_cyto_only_Intensity_IntegratedIntensity_image_GFP',
                                  'obj_final_PI_AreaShape_Area',
                                  'obj_nuclei_Location_Center_X',
                                  'obj_nuclei_Location_Center_Y'), ]
datainput_w <- dcast( variable + location + trackLabel + treatment ~ .id, data = datainput)

datainput_w <- as.data.table(datainput_w)


#source('D://src/lss_fda/ratioDampener.R')

#ratioDampener(dampingFactor = c(0,0.01), nom = datainput[, value],
 #             denom = datainput[, i.value])

#datainput[ , oscivalue := value/ (i.value + 0.005*mean(i.value, na.rm=TRUE))]

unique(datainput$location)

# to test a single track
Frame <- datainput_w[ datainput_w$location == "B05_1" & datainput_w$trackLabel == 12, c("variable")]

value <- datainput_w[ datainput_w$location == "B05_1" & datainput_w$trackLabel == 12, c("obj_cyto_only_Intensity_IntegratedIntensity_image_GFP")]


xCoord <- datainput_w[ datainput_w$location == "B05_1" & datainput_w$trackLabel == 12, c("obj_nuclei_Location_Center_X")]
yCoord <- datainput_w[ datainput_w$location == "B05_1" & datainput_w$trackLabel == 12, c("obj_nuclei_Location_Center_Y")]
signal1 <- datainput_w[ datainput_w$location == "B05_1" & datainput_w$trackLabel == 12, c("obj_final_PI_AreaShape_Area")]




Frame <- Frame$variable # use this for testing a single track
value <- value$obj_cyto_only_Intensity_IntegratedIntensity_image_GFP
xCoord <- xCoord$obj_nuclei_Location_Center_X
yCoord <- yCoord$obj_nuclei_Location_Center_Y
signal1 <- signal1$obj_final_PI_AreaShape_Area


plot(Frame, value)

# use with lapply/ sapply/ vapply over each locationID/trackID combo. input timeID and value


datainput_w[ , uniquelocationtrack:= paste(location, treatment, trackLabel, sep ="_")]
selR <- sample(unique(datainput_w$uniquelocationtrack), 100)

subsetdata <- datainput_w[ datainput_w$uniquelocationtrack %in% selR ,]


subsetdataL <- split(subsetdata, subsetdata$uniquelocationtrack) # turn into list for entries the unique tracks


# and then full dataset..


datainputL <- split(datainput, datainput$uniquelocationtrack)
```


```{r run_analysis }

source('D:/src/lss_fda/celloscillate.R')

# initilly; test the function with a single track;



celloscillate(Frame = Frame, value = value,
              locationID = 'E01_1', timeInterval = '01:34:10',aboveRegr = 5, suppresMaxFrames = 5,
              slopeTH = 0.02,basisType = 'spline',   plotFit = TRUE, plotDomains = TRUE, 
               slopeDomain = NULL, timeToTH = NULL,
              nbasis = 11, lambda = 10, THpeaks = 0.1, f = 0.4,
               signal1 = signal1, signal2 = NULL, signal1TH = 1, xCoord = xCoord, yCoord = yCoord, pix.x = 512, pix.y=512)





C10_1_20


# for callibration first test on subset and set plotting arguments to TRUE
# to ensure you perform a meaningull analysis, select a random subset of around 25-50 tracks of random treatments, optimize parameters, then observe what filter criteria should be used after the analysis to remain with the high quality tracks. 
# make sure tracks are long enough and start early enough

subsetdataL[[2]]

table(unlist(lapply(subsetdataL, function(x) length(x$variable)))) # lengths of time

(lapply(subsetdataL, function(x) (x$variable))) # many tracks start too late


# keep tracks with 40 time points and starting at 1, 2 or 3
indKeep <- unlist(lapply(subsetdataL, function(x) any((x$variable) %in% 1:2)))

subsetdataL <- subsetdataL[ indKeep ]

# keep tracks that are minimaly length 50
indKeep <- unlist(lapply(subsetdataL, function(x) length(x$variable) > 8))
sum(indKeep) # 95 tracks remain
subsetdataL <- subsetdataL[indKeep]


length(subsetdataL)


subsetdataL[[1]]
output <- lapply(subsetdataL, function(x) celloscillate(x, Frame = x$variable, value = x$obj_cyto_only_Intensity_IntegratedIntensity_image_GFP, suppresMaxFrames = 5,
                                            locationID = x$uniquelocationtrack, aboveRegr = 5,timeInterval = '01:06:10',
                                           slopeTH = 0.02, basisType = 'spline',plotFit = TRUE, plotDomains = TRUE, 
                                            slopeDomain = c('00:00:00','06:00:00'), timeToTH = c(0.8, 'relative'),
                           nbasis = 11, lambda = 5, THpeaks = 0.1, f = 0.4,
                           signal1 = x$obj_final_PI_AreaShape_Area, signal2 = NULL, signal1TH = 0.5, pix.x=512, pix.y=512 , xCoord = x$obj_nuclei_Location_Center_X, yCoord = x$obj_nuclei_Location_Center_Y)) 





output <- lapply(datainputL, function(x) celloscillate(x, Frame = x$variable, value = x$oscivalue, suppresMaxFrames = 5,
                                            locationID = x$uniquelocationtrack, aboveRegr = 5,timeInterval = '00:06:10',
                                           slopeTH = 0.02, basisType = 'fourier',plotFit = FALSE, plotDomains = FALSE, 
                                            slopeDomain = c('00:00:00','02:00:00'), timeToTH = c(0.8, 'relative'),
                           nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4,
                           signal1 = x$value, signal2 = x$i.value)) 



# gaps are being closed incorrectly, use domains_F11_1_35
names(output)
names(output[[4]])

# to extract for example all perpeak_peakwidths do
 perpeak_widths <- lapply(output, '[[', 'perpeak_peakwidth' )
# number of peaks ..
 npeaks <- lapply(output, '[[', 'npeaks' )
# etc

 # extract data for certain location by e.g.
 indG08 <- grepl("G08", names(perpeak_widths))

 perpeak_widths[indG08]
# population statistics of parameters e.g. 
 quantile(unlist(perpeak_widths[indG08]))
  

# quaility control: remove too high resid SD, remove non-decreasing peaks, remove if no peaks.
 # also check intensity of cytosol/ nuclei per track and remove based on too low intensity.
 
 

```