---
title: "voorbeeld oscillatie analyse"
author: "Steven Wink"
date: "7 december 2016"
output: html_document
---

 function that performs fourier transform using the fda package toolkit and extracts parameters of interest for in vitro cell biologists
 data input: 
  Frame: integer index of time series
  value: measurement associated with tracked cell that oscillates.
  locationID: string uniquely identifying location and track number of time lapse location and track number, which should hold unique track identifiers you perform the function over.


 plotFit: plot your fits, only do this for a subset during calibration, because writing the pdfs takes alot of time
 nbasis: number of basis functions, increase till smooth fit is obtained. At a certain point no improvement will be noticed.
 lambda: smoothing factor for data. 0 is no smoothing (default max(Frame)/50000) )
 THpeaks: threshold value ratio of pred_peak - lowess(f) and median value, smaller values will not be defined as peaks.   
 f: fraction of data to be used for lowess smoothing, lowess is used for THpeaks. recommend 0.4-0.5. For noisy data increase (noise not id as peak). If lowess line follows fit too closely no maximal/ minima will be defined (also dependent on THpeaks.)
 timeInterval: time between consecutive frames, format "%H:%M:%S"
 aboveRegr : number of points that should be above regression line to identify as peak area)
 basisType : fourier or spline, fourier for osicillatory data and spline for non oscillatory
 slopeDomain : vector of two, first entry the start time - second entry the end time. format "%H:%M:%S. Average slope in this domain is returned
 timeToTH: vector of two, first entry the value, second entry the type: 'relative' or 'absoluut'. returns either the time that the relative threshold value is reached or the time the absoluut value is reached. relative threshold: a fraction between 0 and 1 (0, 1) is defined as the maximum fitted value of the corresponding track * timeToTH. absoluut threshold is defined as the time the timeToTH value is reached
signal1 and signal2: extra data of same length as value and Frame, if provided mean is calculated (usefull for quality control)
slopeTH: regions defined as low slope: first derive below value is defined as low slope. Low slope values are used to determine peak areas (together with local maxima). slopeTH increase will make yellow line gaps larger

the plots: 
a raw fitting plot with residual standard error  
a plot that visualizes the influence of the various tuning paramters. black circles: raw data of track, Blue circle: peak, red circle: valley, yellow dots (zoom in): non-low-slope regions, black dots:regions where fit > lowess regression, black solid-fit line: the model fit, vertical black broken lines: on each side of such a line a defined peak area - but only if it contains a blue circle. These areas are used to determine per - peak measurements. 


```{r data_prep }
require(reshape2)
require(ggplot2)
require(fda)
require(data.table)

# eerst repl2

datainput <- read.delim( file = "repl2/reorderedTrackData.txt", sep ="\t")


datainput$variable <- gsub("TP_", "", datainput$variable)
datainput$variable <- as.numeric(as.character(datainput$variable))

datainput <- datainput[datainput$.id %in% c('obj_shrunkenNuclei_Intensity_IntegratedIntensity_img_GFP', 'obj_cytoN_Intensity_IntegratedIntensity_img_GFP'),]

head(datainput)
cytoN <- datainput[ datainput$.id == 'obj_shrunkenNuclei_Intensity_IntegratedIntensity_img_GFP', ]
shrNuclei <-  datainput[ datainput$.id == 'obj_cytoN_Intensity_IntegratedIntensity_img_GFP', ]

cytoN <- as.data.table(cytoN)
shrNuclei <- as.data.table(shrNuclei)
setkeyv(cytoN, c('location', 'trackLabel', 'variable'))
setkeyv(shrNuclei, c('location', 'trackLabel', 'variable'))
datainput <- cytoN[shrNuclei]


datainput[ , c('mergeLocation', 'i.mergeLocation') := NULL]

source('D://src/lss_fda/ratioDampener.R')

ratioDampener(dampingFactor = c(0,0.01), nom = datainput[, value],
              denom = datainput[, i.value])

datainput[ , oscivalue := value/ (i.value + 0.005*mean(i.value, na.rm=TRUE))]

unique(datainput$location)


Frame <- datainput[ datainput$location == "C10_1" & datainput$trackLabel == 20, c("variable")]
value <- datainput[ datainput$location == "C10_1" & datainput$trackLabel == 20, c("oscivalue")]
Frame <- Frame$variable # use this for testing a single track
value <- value$oscivalue

plot(Frame, value)

# use with lapply/ sapply/ vapply over each locationID/trackID combo. input timeID and value


datainput[ , uniquelocationtrack:= paste(location, trackLabel, sep ="_")]
selR <- sample(unique(datainput$uniquelocationtrack), 100)

subsetdata <- datainput[ datainput$uniquelocationtrack %in% selR ,]
subsetdata$uniquelocationtrack <- paste(subsetdata$location, subsetdata$trackLabel, sep ="_")


subsetdata$location <- as.character(subsetdata$location)
subsetdataL <- split(subsetdata, subsetdata$uniquelocationtrack) # turn into list for entries the unique tracks


# and then full dataset..


datainputL <- split(datainput, datainput$uniquelocationtrack)
```


```{r run_analysis }

source('D:/src/lss_fda/celloscillate.R')

# initilly; test the function with a single track;



celloscillate(Frame = Frame, value = value,
              locationID = 'E01_1', timeInterval = '00:06:10',aboveRegr = 5, suppresMaxFrames = 5,
              slopeTH = 0.02,basisType = 'fourier',   plotFit = TRUE, plotDomains = TRUE, 
               slopeDomain = NULL, timeToTH = NULL,
              nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4,
               signal1 = NULL, signal2 = NULL)
C10_1_20


# for callibration first test on subset and set plotting arguments to TRUE
# to ensure you perform a meaningull analysis, select a random subset of around 25-50 tracks of random treatments, optimize parameters, then observe what filter criteria should be used after the analysis to remain with the high quality tracks. 
# make sure tracks are long enough and start early enough

subsetdataL[[2]]

table(unlist(lapply(subsetdataL, function(x) length(x$variable)))) # lengths of time

(lapply(subsetdataL, function(x) (x$variable))) # many tracks start too late


# keep tracks with 40 time points and starting at 1, 2 or 3
indKeep <- unlist(lapply(subsetdataL, function(x) any((x$variable) %in% 1:3)))
subsetdataL <- subsetdataL[ indKeep ]

# keep tracks that are minimaly length 50
indKeep <- unlist(lapply(subsetdataL, function(x) length(x$variable) > 49))
sum(indKeep) # only 16 % of tracks remain
subsetdataL <- subsetdataL[indKeep]
length(subsetdataL)
output <- lapply(subsetdataL, function(x) celloscillate(x, Frame = x$variable, value = x$oscivalue, suppresMaxFrames = 5,
                                            locationID = x$uniquelocationtrack, aboveRegr = 5,timeInterval = '00:06:10',
                                           slopeTH = 0.02, basisType = 'fourier',plotFit = TRUE, plotDomains = TRUE, 
                                            slopeDomain = c('00:00:00','02:00:00'), timeToTH = c(0.8, 'relative'),
                           nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4,
                           signal1 = x$value, signal2 = x$i.value)) 



lapply(datainputL

output <- lapply(datainputL, function(x) celloscillate(x, Frame = x$variable, value = x$oscivalue, suppresMaxFrames = 5,
                                            locationID = x$uniquelocationtrack, aboveRegr = 5,timeInterval = '00:06:10',
                                           slopeTH = 0.02, basisType = 'fourier',plotFit = FALSE, plotDomains = FALSE, 
                                            slopeDomain = c('00:00:00','02:00:00'), timeToTH = c(0.8, 'relative'),
                           nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4,
                           signal1 = x$value, signal2 = x$i.value)) 



# gaps are being closed incorrectly, use domains_F11_1_35
names(output)
names(output[[4]])

# to extract for example all perpeak_peakwidths do
 perpeak_widths <- lapply(output, '[[', 'perpeak_peakwidth' )
# number of peaks ..
 npeaks <- lapply(output, '[[', 'npeaks' )
# etc

 # extract data for certain location by e.g.
 indG08 <- grepl("G08", names(perpeak_widths))

 perpeak_widths[indG08]
# population statistics of parameters e.g. 
 quantile(unlist(perpeak_widths[indG08]))
  

# quaility control: remove too high resid SD, remove non-decreasing peaks, remove if no peaks.
 # also check intensity of cytosol/ nuclei per track and remove based on too low intensity.
 
 

```