---
title: "voorbeeld oscillatie analyse"
author: "Steven Wink"
date: "7 december 2016"
output: html_document
---

 function that performs fourier transform using the fda package toolkit and extracts parameters of interest for in vitro cell biologists
 data input: 
  Frame: integer index of time series
  value: measurement associated with tracked cell that oscillates.
  locationID: name of time lapse location, which should hold unique track identifiers you perform the function over.


 plotFit: plot your fits, only do this for a subset during calibration, because writing the pdfs takes alot of time
 nbasis: number of basis functions, increase till smooth fit is obtained. At a certain point no improvement will be noticed.
 lambda: smoothing factor for data. 0 is no smoothing
 THpeaks: threshold value ratio of pred_peak - lowess(f) and median value, smaller values will not be defined as peaks.   
 f: fraction of data to be used for lowess smoothing, lowess is used for THpeaks. recommend 0.4-0.5
 timeInterval: time between consecutive frames, format "%H:%M:%S"
 aboveRegr : number of points that should be above regression line to identify as peak area)

```{r data_prep }
require(reshape2)
require(ggplot2)
require(fda)
require(data.table)

datainput <- read.delim( file = "../data/reorderedTrackData.txt", sep ="\t")
head(datainput)

datainput$variable <- gsub("TP_", "", datainput$variable)
datainput$variable <- as.numeric(as.character(datainput$variable))

datainput <- datainput[datainput$.id %in% c('obj_shrunkenNuclei_Intensity_IntegratedIntensity_img_GFP', 'obj_cytoN_Intensity_IntegratedIntensity_img_GFP'),]

head(datainput)
cytoN <- datainput[ datainput$.id == 'obj_shrunkenNuclei_Intensity_IntegratedIntensity_img_GFP', ]
shrNuclei <-  datainput[ datainput$.id == 'obj_cytoN_Intensity_IntegratedIntensity_img_GFP', ]

cytoN <- as.data.table(cytoN)
shrNuclei <- as.data.table(shrNuclei)
setkeyv(cytoN, c('location', 'trackLabel', 'variable'))
setkeyv(shrNuclei, c('location', 'trackLabel', 'variable'))
datainput <- cytoN[shrNuclei]
gc()

datainput[ , c('mergeLocation', 'i.mergeLocation') := NULL]

#ratioDampener(dampingFactor = c(0,0.1), nom = datainput[, value],
#              denom = datainput[, i.value], ylim = c(0,15))

datainput[ , oscivalue := value/ (i.value + 0.05*mean(i.value, na.rm=TRUE))]


unique(datainput$location)

Frame <- datainput[ datainput$location == "E02_1" & datainput$trackLabel == 8, c("variable")]
value <- datainput[ datainput$location == "E02_1" & datainput$trackLabel == 8, c("oscivalue")]
Frame <- Frame$variable # use this for testing a single track
value <- value$oscivalue

plot(Frame, value)

# use with lapply/ sapply/ vapply over each locationID/trackID combo. input timeID and value
subsetdata <- datainput[ datainput$location %in% c("E02_1", "C03_1", "G08_1", "D06_1") & datainput$trackLabel %in% 5:8,]
subsetdata$uniquelocationtrack <- paste(subsetdata$location, subsetdata$trackLabel, sep ="_")
subsetdata$location <- as.character(subsetdata$location)
subsetdataL <- split(subsetdata, subsetdata$uniquelocationtrack) # turn into list for entries the unique tracks


# and then full dataset..


```


```{r run_analysis }

source('celloscillate.R')

# initilly; test the function with a single track;



celloscillate(Frame = Frame, value = value,
                                            locationID = 'E01_1', timeInterval = '00:06:10',aboveRegr = 5,
                                            plotFit = TRUE, plotDomains = TRUE, 
                           nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4)



# for callibration first test on subset and set plotting arguments to TRUE



output <- lapply(subsetdataL, function(x) celloscillate(x, Frame = x$variable, value = x$oscivalue, 
                                            locationID = x$location, aboveRegr = 5,timeInterval = '00:06:10',
                                            plotFit = TRUE, plotDomains = TRUE, 
                           nbasis = 19, lambda = NULL, THpeaks = 0.1, f = 0.4)) 


names(output)
names(output[[4]])

# to extract for example all perpeak_peakwidths do
 perpeak_widths <- lapply(output, '[[', 'perpeak_peakwidth' )
# number of peaks ..
 npeaks <- lapply(output, '[[', 'npeaks' )
# etc

 # extract data for certain location by e.g.
 indG08 <- grepl("G08", names(perpeak_widths))

 perpeak_widths[indG08]
# population statistics of parameters e.g. 
 quantile(unlist(perpeak_widths[indG08]))
  

# quaility control: remove too high resid SD, remove non-decreasing peaks, remove if no peaks.
 # also check intensity of cytosol/ nuclei per track and remove based on too low intensity.
 
 

```